---
title: "Chapter 3"
author: "Wenduo Wang"
date: "July 20, 2016"
output: html_document
---

Problem  15

(a)
Let's create a simple linear regression model for each variable against `crim`, and plot the fitted line and actual data points.
```{r}
library(MASS)
library(dplyr)
summary(Boston)
row_no <- nrow(Boston)
training_rows <- sample(row_no, round(row_no*0.8))
training_set <- Boston[training_rows,]
test_set <- setdiff(Boston, training_set)
confint_df <- data.frame()
par(mfrow=c(4, 4), mar=c(2, 1, 2, 1))
for (i in c(1:ncol(Boston))){
    if (colnames(Boston)[i]=="crim"){
        next
    }
    formula <- as.formula(paste("crim~", colnames(Boston)[i]))
    lm_model <- lm(formula, data=training_set)
    plot(y=Boston$crim, x=Boston[,i], main=paste("crim ~", colnames(Boston)[i]), pch=1, col="lightgray")
    abline(lm_model$coef[1], lm_model$coef[2], col="red")
    confint_df <- rbind(confint_df, confint(lm_model)[2,])
}
print(dim(confint_df))
colnames(confint_df) <- c("2.5%", "97.5%")
row.names(confint_df) <- colnames(Boston)[colnames(Boston)!="crim"]
```

The linear model does not fit the data very well according to the plots. To further assess the correlation, let's have a closer look at the coefficients of these models

```{r}
print(confint_df)
```

The above code lists the 95% confindence intervals of coefficients for all the predictors with respect to each linear model. As seen from the result, the confident interval of the coefficient of `chas` contains zero, which means this predictor is probably not correlated with `crim`. Meanwhile, other predictors mostly have very small coefficients close to 0, with the exception of `nox`. So up to now, it appears `nox` is most likely to be a true predictor of `crim`.

(b, c)
Now let's create a linear model of `crim` on all predictors, and see their coefficients' confident intervals.
```{r}
lm_model_multiple <- lm(crim~., data=training_set)
print(confint(lm_model_multiple))
```

When all the predictors are simultaneously fitted against `crim`, the result turns out very different from the previous single linear regression models.
First, the previously strong predictor `nox` is now largely irrelevant. Actually in this case only `zn`, `dis` and `rad` show a significant correlation with `crim`, which is equivalent to rejecting the null hypothesis that $$\beta_[0]$$